---
title: "US_Coasters_By_The_Numbers"
author: "Austin Wright"
date: "1/14/2022"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(plotly)
library(htmlwidgets)
library(plyr)
library(reshape2)
library(data.table)
library(compositions)
#knitr::opts_chunk$set(echo = FALSE)
options(warn=-1)
```

## R Markdown

This is an R Markdown Test of my new data set from RCDB. 


```{r data_set}
Stats_2021 <- read.csv("./data/US_Coaster_Stats_2021.csv")
```

## Intro
Using RCDB's data

You can also embed plots, for example:

```{r Population Percentage}
column_name = c("State","Percent")
topStates = table(Stats_2021$State)
topStates = as.data.frame(topStates)
colnames(topStates)<-column_name
topStates = top_n(topStates,10,Percent)
topSum <- sum(topStates$Percent)

#Get the percentages of each state
for (state in topStates$Percent){
  percentages <- round(((topStates$Percent/(topSum))*100.0),digits=0)
}

piePlot <- ggplot(topStates, aes(x = "", y = Percent, fill = State)) +
  geom_col() + coord_polar(theta = "y", start=0) +  
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid  = element_blank()) + 
  geom_text(aes(label = paste0((State)," ",(percentages), "%")),
        position = position_stack(vjust = 0.5), angle = c(-60,-13,20,50,70,-80,-50,-10,25,70)) + 
  ggtitle("Top 10 States With the Most Coasters")

piePlot
```


1.What are the most popular manufactures? 
```{r Top 10 Most Popular Manufactures}
column_name = c("Manufacture", "Installations")
topManufactures = table(Stats_2021$Make)
topManufactures = as.data.frame(topManufactures)
colnames(topManufactures) <- column_name
topManufactures = top_n(topManufactures, 10, Installations)
barplot <- ggplot(data = topManufactures, width = 700,height = 700, mapping = aes(x = Manufacture, y = Installations)) + 
  geom_bar(stat="identity", fill=rainbow(10),position=position_dodge()) + 
  geom_text(aes(label = paste0((Installations), " installs")), position = position_stack(vjust = 0.5)) + 
  #theme(axis.text.x=element_blank(),
        #axis.ticks.x=element_blank()) +
  coord_flip() + 
  ggtitle("Top 10 Most Popular Manufactures")
ggplotly(barplot, tooltip = "text")


```

Random Variables of the data set that show the history of coasters.
The End of the 19th century...

```{r  Population }

WestStates = c("California","Nevada","Arizona","Idaho", "Oregon","Washington", "New Mexico","Colorado" , "Utah", "Montana", "Wyoming")
EastStates = c("Maine" , "Massachusetts", "Vermont" ,"New Hampshire" ,"Rhode Island","Connecticut", "New York" ,"Delaware" , "Maryland" , "Virginia" , "Pennsylvania" , "New Jersey" ,"Virginia" , "North Carolina" , "South Carolina" , "Georgia" , "Florida")
Makers = c("Intamin Amusement Rides", "Bolliger & Mabillard")

westCoastIntamin <- nrow(data_frame(Stats_2021) %>% select("Name","State","Make") %>% filter(State %in% WestStates) %>% filter(Make == Makers[1]))
eastCoastIntamin <- nrow(data_frame(Stats_2021) %>% select("Name","State","Make") %>% filter(State %in% EastStates) %>% filter(Make == Makers[1]))
westCoastBM <- nrow(data_frame(Stats_2021) %>% select("Name","State","Make") %>% filter(State %in% WestStates) %>% filter(Make == Makers[2]))
eastCoastBM <-  nrow(data_frame(Stats_2021) %>% select("Name","State","Make") %>% filter(State %in% EastStates) %>% filter(Make == Makers[2]))

maker <- c(rep("Intamin", 2), rep("Bollinger and Mabillard",2))
region <- rep(c("West", "East"),2)
value <- c(westCoastIntamin,eastCoastIntamin,westCoastBM,eastCoastBM)
regionPopulations <- data.frame(maker,region,value)

barplot <- ggplot(regionPopulations,width = 700,height = 700,aes(fill=region,y=value,x=maker)) + geom_bar(position="stack",stat="identity") +
   geom_text(aes(label = paste0((value),  " installs")), position = position_stack(vjust = 0.5)) + 
  coord_flip() + 
  ggtitle("Swiss Maker Installs Between Coasts")
ggplotly(barplot, tooltip = "text")


```
#Probability of a coaster being made by Intamin given it is on the west coast?
#A: Made by Intamin
#B. Located on the west coast
#P(A | B) = P(A intersect B) / P(B)
#P(B) = Total
```{r Prob Intamin}
probTotal <- westCoastIntamin + eastCoastIntamin + westCoastBM + eastCoastBM
probWestIntamin <- (westCoastIntamin)/(probTotal)
probEastIntamin <- (eastCoastIntamin)/(probTotal)

probTotal <- westCoastIntamin + eastCoastIntamin + westCoastBM + eastCoastBM
probWestBM <- (westCoastBM)/(probTotal)
probEastBM <- (eastCoastBM)/(probTotal)

maker <- c(rep("Intamin", 2), rep("Bollinger and Mabillard",2))
region <- rep(c("West", "East"),2)
value <- c(probWestIntamin,probEastIntamin,probWestBM,probEastBM)
regionPopulations <- data.frame(region,value)

barplot <- ggplot(regionPopulations,width = 700,height = 700,aes(fill=region,y=value,x=maker)) + geom_bar(position="stack",stat="identity") +
   geom_text(aes(label = paste0((value))), position = position_stack(vjust = 0.5)) + 
  ggtitle("Probability Of Swiss Installs Between Coasts")

ggplotly(barplot, tooltip = "text")
```


#Standard Histogram of height
```{r Height}
stats <- as.data.frame(Stats_2021) %>% select(Height,Material)
stats <- na.omit(stats)

sample_steel <- stats %>% filter(Material=="Steel")
sample_wood  <- stats %>% filter(Material=="Wood")

mean_steel <- round(mean(sample_steel$Height))
mean_wood <- round(mean(sample_wood$Height))
std_steel <- round(sd(sample_steel$Height))
std_wood <- round(sd(sample_wood$Height))
sample_size_wood <- nrow(sample_wood)
sample_size_steel <- nrow(sample_steel)
sample_size <- nrow(stats)
stats <- data.frame(
  Material=factor(rep(c("Steel", "Wood"),each=sample_size)),
  Height=round(c(rnorm(sample_size_steel, mean=mean_steel, sd=std_steel), rnorm(sample_size_wood, mean=mean_wood, sd=std_wood)))
)

ggplot(stats,aes(y=Height,fill=Material)) + geom_histogram() + 
  ggtitle("Normal Distribution of US Coaster Height") 


```


#Standard Histogram of height
```{r Inversions}
stats <- as.data.frame(Stats_2021) %>% select(Inversions,Material)
stats <- na.omit(stats)

sample_steel <- stats %>% filter(Material=="Steel")
sample_wood  <- stats %>% filter(Material=="Wood")

mean_steel <- round(mean(sample_steel$Inversions))
mean_wood <- round(mean(sample_wood$Inversions))
std_steel <- round(sd(sample_steel$Inversions))
std_wood <- round(sd(sample_wood$Inversions))

sample_size_wood <- nrow(sample_wood)
sample_size_steel <- nrow(sample_steel)
sample_size <- nrow(stats)

stats <- data.frame(
  Material=factor(rep(c("Steel", "Wood"),each=sample_size)),
  Inversions=round(c(rnorm(sample_size_steel, mean=mean_steel, sd=std_steel), rnorm(sample_size_wood, mean=mean_wood, sd=std_wood)))
)

ggplot(stats,aes(y=Inversions,fill=Material)) + geom_histogram() + ggtitle("Normal Distribution of US Coaster Inversions")

```
#T Distribution Histogram of height
```{r Height2}
stats <- as.data.frame(Stats_2021) %>% select(Inversions,Material)
stats <- na.omit(stats)

sample_steel <- stats %>% filter(Material=="Steel")
sample_wood  <- stats %>% filter(Material=="Wood")

mean_steel <- round(mean(sample_steel$Inversions))
mean_wood <- round(mean(sample_wood$Inversions))
std_steel <- round(sd(sample_steel$Inversions))
std_wood <- round(sd(sample_wood$Inversions))
sample_size <- nrow(stats)

stats <- data.frame(
  Material=factor(rep(c("Steel", "Wood"),each=sample_size)),
  Inversions=round(c(rnorm(sample_size, mean=mean_steel, sd=std_steel), rnorm(sample_size, mean=mean_wood, sd=std_wood)))
)

ggplot(stats,aes(y=Inversions,fill=Material)) + geom_histogram() + ggtitle("Normal Distribution of US Coaster Inversions")

```


